Trauma treatment delay

### 2023-12-12 manuscriptlongversion.Rmd
text

These results show a possible correlation between some variables and the
presence of delayed treatment. A lower GCS had a correlation with
delayed treatment while a higher ISS had the same correlation. This
suggests that a more heavily injured and less awake patient is more
prone to receiving delayed treatment. Compared to previous a previous
study analysing the same data but with ofi as the outcome, this result
is similar to what they found using the same database [@Pat_factors]. As
the mentioned study suggests, this could be explained by that lower GCS
indicates severe trauma, aswell as higher ISS. More injured patients
often entails a more complex management and require more interventions.
More interventions leave more room for delay to these interventions, why
they are more prone to delay in treatment.

Respiratory rate also had the correlation that lower respiratory rate
correlated with lower rates of delayed treatment among the patients. As
the interval is \>10 to \<29 in the triage guideline, both too low and
too high respiratory rate is not desirable[@EHCP2015]. This might have a
similar explanation to ISS and GCS, that patients with tachypnea might
receive higher rates of delayed treatment because they are more severely
injured, therefore a higher risk of missing a possible injury. Another
possible explanation might be bradypnea in a patient might attract more
attention in the hospital and lower the incidence of delayed treatment.
Another similar study, which focused on ofi generally, showed the same
correlation, although the significance dissapeared in the adjusted
analysis [@Pat_factors].

Number of injuries was significant when unadjusted, but the correlation
dissapeared when adjusted for with other variables. Our guess as to why
this is due to confounding due to it's shared variability with the ISS
variable. Number of injuries and ISS is often correlated as a more
severely injured patient will both score higher on ISS and number of
injuries. Therefore, we hypothesize that the effect of number of
injuries become insignificant when adjusting for ISS.

When looking at the patient demographic, it is clear that the majority
admitted to trauma care are men, with it being slightly higher in the
delay to treatment group. The fact that age and gender did not show an
increased odds ratio should be seen as a positive sign. It suggests that
among trauma patients, there is no systematic delay based on gender or
age.

>>>>

These results show a possible correlation between some variables and the
presence of delayed treatment. A lower GCS had a correlation with
delayed treatment while a higher ISS had the same correlation. This
suggests that a more heavily injured and less awake patient is more
prone to receiving delayed treatment. Compared to previous a previous
study analysing the same data but with ofi as the outcome, this result
is similar to what they found using the same database [@Pat_factors]. As
the mentioned study suggests, this could be explained by that lower GCS
indicates severe trauma, as well as higher ISS. More injured patients
often entails a more complex management and require more interventions.
More interventions leave more room for delay to these interventions, why
they are more prone to delay in treatment. Respiratory rate also had the
correlation that lower respiratory rate correlated with lower rates of
delayed treatment among the patients. As the interval is \>10 to \<29 in
the triage guideline, both too low and too high respiratory rate is not
desirable[@EHCP2015]. This might have a similar explanation to ISS and
GCS, that patients with tachypnea might receive higher rates of delayed
treatment because they are more severely injured, therefore a higher
risk of missing a possible injury. A previous study also found this, but
the correlation dissapeared when adjusted for, which was hypothesised
being due to it's shared variability with the ISS
variable[@Pat_factors]. In this study, this was the case for number of
injuries, as it and ISS is often correlated as a more severely injured
patient will both score higher on ISS and number of injuries. Therefore,
we hypothesize that the effect of number of injuries become
insignificant when adjusting for ISS.

Not many studies regarding trauma and patient level factors have had
delayed treatment as their outcome, with many analysing mortality. Some
likeness exist with which patient level factors predict the outcome,
with higher mortality being seen in patients with lower GCS and higher
ISS [@Mort_sam][@Mort_hef]. Increased age was also a factor linked to
mortality, which was not seen to be linked to delayed treatment in this
study. This

When looking at the patient demographic, it is clear that the majority
admitted to trauma care are men, with it being slightly higher in the
delay to treatment group. However, not high enough to be statistically
significant. The fact that gender did not show an increased odds ratio
should be seen as a positive sign. It suggests that among trauma
patients, there is no systematic delay based on this patient level
factor. The same goes for age, as
